### 2022-04-29 dev/submissions/aies/extended_abstract/extended_abstract.qmd
first complete draft

Recent advances in artificial intelligence (AI) have propelled its adoption in domains outside of computer science including health care, bioinformatics and genetics. In finance, economics and other social sciences, applications of AI are still relatively limited. Decision-making in these fields has traditionally been guided by interpretable models that facilitate explanations. Explainability is crucial in the social sciences context, because inference is typically at least as important as predictive performance. Decision-makers in the social sciences are also typically required to explain their decisions to human stakeholders: central bankers, for example, are held accountable by the public for the policies they decide on. It is therefore not surprising that practitioners and academics in these fields are reluctant to adopt AI technologies that cannot be explained. Deep neural networks, for example, are generally considered as black boxes and therefore not trustworthy in a context that demands explanations. This PhD project is focused on exploring and developing methodologies that improve the trustworthiness of AI and thereby enable its application in Finance and Economics.

The remainder of this extended abstract is structured as follows: @sec-main presents one of the research questions I have investigated during the first months of my PhD: how do counterfactual explanations handle dynamics? @sec-related places this work in the broader context of Trustworthy AI for Finance and Economics.

@@

**Counterfactual explanations** (CE) explain how inputs into a model need to change for it to produce different outputs. They are intuitive, simple and intrinsically linked to the potential outcome framework for causal inference, which social scientists are familiar with. Counterfactual explanations that involve realistic and actionable changes can be used for the purpose of **Algorithmic Recourse** (AR) to help individuals facing adverse decisions. An example relevant to the Finance and Economics domain is consumer credit: in this context AR can be used to guide individuals to improve their credit worthiness, should they have previously been denied access to credit based on a black-box decision-making system .

Existing work on CE and AR has largely been limited to the static setting: given some classifier $M: \mathcal{X} \mapsto \mathcal{Y}$ we are interested in finding close [@wachter2017counterfactual], actionable [@ustun2019actionable], realistic [@joshi2019towards, @schut2021generating], sparse, diverse [@mothilal2020explaining] and ideally causally founded counterfactual explanations [@karimi2021algorithmic] for some individual $x$. The ability of counterfactual explanations to handle dynamics like data and model shifts remains a largely unexplored research challenge at this point [@verma2020counterfactual]. Only one recent work considers the implications of **exogenous** domain shifts on the validity of recourse [@upadhyay2021towards]. The authors propose a simple minimax objective, that minimizes the counterfactual loss function for a maximal domain and model shift. They show that their approach yields more robust counterfactuals than existing approaches.

This project investigates **endogenous** domain and model shifts, that is shifts that occur when AR is actually implemented by a proportion of individuals and the classifier is updated in response. @fig-dynamics illustrates this idea for a binary problem involving a probabilistic classifier and a greedy counterfactual generator proposed by @schut2021generating: AR leads to a domain shift, which in turn causes a drastic model shift. As this game of implementing AR and updating the classifier is repeated, the decision boundary moves in the opposite direction of the original training samples in the target class. We consider several aspects of these dynamics as problematic. Firstly, as the decision boundary moves in the direction of the non-target class, counterfactual paths become shorter: in the loan example, individuals that previously would have been denied credit based on their input features are suddenly considered as credit worthy. Average default risk across all borrowers can therefore be expected to increase. Conversely, lenders that anticipate or detect these sorts of dynamics may choose to still deny credit to individuals that have implemented AR, thereby comprimising the validity of AR.

To the best of my knowledge this is the first work investigating endogenous dynamic in algorithmic recourse. In future experiments I want to investigate how this phenomenon plays out across different benchmark datasets including German credit, Boston Housing and COMPAS. Furthermore, I want to assess to what extent the magnitude and direction of domain and model shifts depends on the choice of the counterfactual generator. To this end, I am currently supervising a group of undergraduate students, who are tackling some of these tasks in their final-year research project.

![PLACEHOLDER: The dynamics of algorithmic recourse.](){#fig-dynamics fig.pos="h" height=100px}

@@

Until recently there existed only one open-source library that provides a unifying approach to generate and benchmark counterfactual explanations for models built and trained in Python [@pawelczyk2021carla]. To address this limitation I have developed [CounterfactualExplanations.jl](https://www.paltmeyer.com/CounterfactualExplanations.jl/stable/): a Julia package that can be used to generate counterfactual explanations for models developed and trained not only in Julia, but also in other popular programming languages like Python and R. The package and companion paper are currently pending acceptance for a main talk at [JuliaCon](https://juliacon.org/2022/).

@@

To ensure that the generated explanations are realistic it is important to understand which input-output pairs are likely and which are not. With respect to the work-in-progress presented here, I expect that this may help in mitigating endogenous domain and model shifts. To quantify the joint likelihood of input-output pairs, previous work has either relied on generative models or restricted the analysis to probabilistic models that incorporate uncertainty in their predictions. While the former approach is more versatile, since it is applicable to both deterministic and probabilistic models, the latter is computationally much more efficient. The approach proposed by @schut2021generating and used to generate the examples in @fig-dynamics falls into the latter category.  In future work, I want to explore how recent advances in post-hoc uncertainty quantification, most notably Laplace Redux [@daxberger2021laplace], can be leveraged to generate realistic and unambiguous counterfactual explanations for any model.^[For some initial work on this see my Julia implementation of Laplace Redux: [BayesLaplace.jl](https://www.paltmeyer.com/BayesLaplace.jl/dev/).]

>>>>

Recent advances in artificial intelligence (AI) have propelled its adoption in domains outside of computer science including health care, bioinformatics and genetics. In finance, economics and other social sciences, applications of AI are still relatively limited. Decision-making in these fields has traditionally been guided by interpretable models that facilitate explanations. Explainability is crucial in this context, since decision-makers are typically held accountable by the public: central banks, for example, are heavily scrutinized for the policies they impose. It is therefore not surprising that practitioners and academics in these fields are reluctant to adopt AI technologies they cannot trust. Deep neural networks, for example, are generally considered as black boxes and therefore not trustworthy in a context that demands explanations. This PhD project is focused on exploring and developing methodologies that improve the trustworthiness of AI and thereby enable its application in Finance and Economics.

The remainder of this extended abstract is structured as follows: @sec-main presents one of the research questions I have investigated during the first months of my PhD: how do counterfactual explanations handle dynamics? @sec-related places this work in the broader context of my research.

@@

**Counterfactual explanations** (CE) explain how inputs into a model need to change for it to produce different outputs. They are intuitive, simple and intrinsically linked to the potential outcome framework for causal inference, which social scientists are familiar with. Counterfactual explanations that involve realistic and actionable changes can be used for the purpose of **Algorithmic Recourse** (AR) to help individuals who face adverse outcomes. An example relevant to the Finance and Economics domain is consumer credit: in this context AR can be used to guide individuals in improving their creditworthiness, should they have previously been denied access to credit based on an automated decision-making system.

Existing work on CE and AR has largely been limited to the static setting: given some classifier $M: \mathcal{X} \mapsto \mathcal{Y}$ we are interested in finding close [@wachter2017counterfactual], actionable [@ustun2019actionable], realistic [@joshi2019towards, @schut2021generating], sparse, diverse [@mothilal2020explaining] and ideally causally founded counterfactual explanations [@karimi2021algorithmic] for some individual $x$. The ability of counterfactual explanations to handle dynamics like data and model shifts remains a largely unexplored research challenge at this point [@verma2020counterfactual]. Only one recent work considers the implications of **exogenous** domain and model shifts [@upadhyay2021towards]. The authors propose a simple minimax objective, that minimizes the counterfactual loss function for a maximal model shift. They show that their approach yields more robust counterfactuals in this context than existing approaches.

This project investigates **endogenous** domain and model shifts, that is shifts that occur when AR is actually implemented by a proportion of individuals and the classifier is updated in response. @fig-dynamics illustrates this idea for a binary problem involving a probabilistic classifier and the counterfactual generator proposed by @wachter2017counterfactual: the implementation of AR for a subset of individuals leads to a domain shift (b), which in turn triggers a model shift (c). As this game of implementing AR and updating the classifier is repeated, the decision boundary moves away from training samples that were originally in the target class (d).

These dynamics may be problematic. As the decision boundary moves in the direction of the non-target class, counterfactual paths become shorter: in the loan example, individuals that previously would have been denied credit based on their input features are suddenly considered as creditworthy. Average default risk across all borrowers can therefore be expected to increase. Conversely, lenders that anticipate such dynamics may choose to deny credit to individuals that have implemented AR, thereby compromising the validity of AR.

To the best of my knowledge this is the first work investigating endogenous dynamic in AR. Through future experiments I want to investigate how this phenomenon plays out across different benchmark datasets including German credit, Boston Housing and COMPAS.^[These benchmark datasets have their issues and controversies, which is one of the challenges I would like to discuss at AIES.] Furthermore, I want to assess to what extent the magnitude and direction of domain and model shifts depends on the choice of the counterfactual generator. To this end, I am currently supervising a group of undergraduate students, who are tackling some of these tasks in their final-year research project.

![Dynamics in Algorithmic Recourse: we have a simple Bayesian model trained for binary classification (a); the implementation of AR for a random subset of individuals leads to a domain shift (b); as the classifier is retrained we observe a model shift (c); as this process is repeated, the decision boundary moves away from the target class (d).](www/poc.png){#fig-dynamics fig.pos="h" width=45%}

@@

Until recently there existed only one open-source library that provides a unifying approach to generate and benchmark counterfactual explanations for Python models [@pawelczyk2021carla]. To address this limitation I have developed [CounterfactualExplanations.jl](https://www.paltmeyer.com/CounterfactualExplanations.jl/stable/): a Julia package that can be used to generate counterfactual explanations for models developed and trained not only in Julia, but also in other popular programming languages. The package and companion paper are pending acceptance for a main talk at [JuliaCon '22](https://juliacon.org/2022/).

@@

To ensure that the generated counterfactuals are realistic it helps to understand which input-output pairs are likely to occur under the data generating process. To this end, previous work has either relied on generative models or restricted the analysis to probabilistic classifiers that incorporate uncertainty in their predictions. While the former approach is more generally applicable, the latter is computationally more efficient. In future work, I want to explore how recent advances in post-hoc uncertainty quantification, most notably Laplace Redux [@daxberger2021laplace], can be leveraged to generate realistic and unambiguous counterfactual explanations for any model.^[For some initial work on this see my Julia implementation of Laplace Redux: [BayesLaplace.jl](https://www.paltmeyer.com/BayesLaplace.jl/dev/).] With respect to the work-in-progress presented here, I expect that these efforts may help in mitigating endogenous domain and model shifts.

### 2022-04-29 dev/submissions/aies/extended_abstract/extended_abstract.tex
first complete draft

\title{Dynamics in Algorithmic Recourse}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Trustworthy Artificial Intelligence for Finance and Economics}
\author{}
\date{April 26, 2022}

@@

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, frame hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, sharp corners, interior hidden, boxrule=0pt]}{\end{tcolorbox}}\fi

@@

Recent advances in artificial intelligence (AI) have propelled its
adoption in domains outside of computer science including health care,
bioinformatics and genetics. In finance, economics and other social
sciences, applications of AI are still relatively limited.
Decision-making in these fields has traditionally been guided by
interpretable models that facilitate explanations. Explainability is
crucial in the social sciences context, because inference is typically
at least as important as predictive performance. Decision-makers in the
social sciences are also typically required to explain their decisions
to human stakeholders: central bankers, for example, are held
accountable by the public for the policies they decide on. It is
therefore not surprising that practitioners and academics in these
fields are reluctant to adopt AI technologies that cannot be explained.
Deep neural networks, for example, are generally considered as black
boxes and therefore not trustworthy in a context that demands
explanations. This PhD project is focused on exploring and developing
methodologies that improve the trustworthiness of AI and thereby enable
its application in Finance and Economics.

The remainder of this extended abstract is structured as follows:
Section~\ref{sec-main} presents one of the research questions I have
investigated during the first months of my PhD: how do counterfactual
explanations handle dynamics? Section~\ref{sec-related} places this work
in the broader context of Trustworthy AI for Finance and Economics.

@@

\textbf{Counterfactual explanations} (CE) explain how inputs into a
model need to change for it to produce different outputs. They are
intuitive, simple and intrinsically linked to the potential outcome
framework for causal inference, which social scientists are familiar
with. Counterfactual explanations that involve realistic and actionable
changes can be used for the purpose of \textbf{Algorithmic Recourse}
(AR) to help individuals facing adverse decisions. An example relevant
to the Finance and Economics domain is consumer credit: in this context
AR can be used to guide individuals to improve their credit worthiness,
should they have previously been denied access to credit based on a
black-box decision-making system .

Existing work on CE and AR has largely been limited to the static
setting: given some classifier \(M: \mathcal{X} \mapsto \mathcal{Y}\) we
are interested in finding close (Wachter, Mittelstadt, and Russell
2017), actionable (Ustun, Spangher, and Liu 2019), realistic Schut et
al. (2021), sparse, diverse (Mothilal, Sharma, and Tan 2020) and ideally
causally founded counterfactual explanations (Karimi, Schölkopf, and
Valera 2021) for some individual \(x\). The ability of counterfactual
explanations to handle dynamics like data and model shifts remains a
largely unexplored research challenge at this point (Verma, Dickerson,
and Hines 2020). Only one recent work considers the implications of
\textbf{exogenous} domain shifts on the validity of recourse (Upadhyay,
Joshi, and Lakkaraju 2021). The authors propose a simple minimax
objective, that minimizes the counterfactual loss function for a maximal
domain and model shift. They show that their approach yields more robust
counterfactuals than existing approaches.

This project investigates \textbf{endogenous} domain and model shifts,
that is shifts that occur when AR is actually implemented by a
proportion of individuals and the classifier is updated in response.
Figure~\ref{fig-dynamics} illustrates this idea for a binary problem
involving a probabilistic classifier and a greedy counterfactual
generator proposed by Schut et al. (2021): AR leads to a domain shift,
which in turn causes a drastic model shift. As this game of implementing
AR and updating the classifier is repeated, the decision boundary moves
in the opposite direction of the original training samples in the target
class. We consider several aspects of these dynamics as problematic.
Firstly, as the decision boundary moves in the direction of the
non-target class, counterfactual paths become shorter: in the loan
example, individuals that previously would have been denied credit based
on their input features are suddenly considered as credit worthy.
Average default risk across all borrowers can therefore be expected to
increase. Conversely, lenders that anticipate or detect these sorts of
dynamics may choose to still deny credit to individuals that have
implemented AR, thereby comprimising the validity of AR.

To the best of my knowledge this is the first work investigating
endogenous dynamic in algorithmic recourse. In future experiments I want
to investigate how this phenomenon plays out across different benchmark
datasets including German credit, Boston Housing and COMPAS.
Furthermore, I want to assess to what extent the magnitude and direction
of domain and model shifts depends on the choice of the counterfactual
generator. To this end, I am currently supervising a group of
undergraduate students, who are tackling some of these tasks in their
final-year research project.

@@

\caption{\label{fig-dynamics}PLACEHOLDER: The dynamics of algorithmic
recourse.}

@@

Until recently there existed only one open-source library that provides
a unifying approach to generate and benchmark counterfactual
explanations for models built and trained in Python (Pawelczyk et al.
2021). To address this limitation I have developed
\href{https://www.paltmeyer.com/CounterfactualExplanations.jl/stable/}{CounterfactualExplanations.jl}:
a Julia package that can be used to generate counterfactual explanations
for models developed and trained not only in Julia, but also in other
popular programming languages like Python and R. The package and
companion paper are currently pending acceptance for a main talk at
\href{https://juliacon.org/2022/}{JuliaCon}.

@@

To ensure that the generated explanations are realistic it is important
to understand which input-output pairs are likely and which are not.
With respect to the work-in-progress presented here, I expect that this
may help in mitigating endogenous domain and model shifts. To quantify
the joint likelihood of input-output pairs, previous work has either
relied on generative models or restricted the analysis to probabilistic
models that incorporate uncertainty in their predictions. While the
former approach is more versatile, since it is applicable to both
deterministic and probabilistic models, the latter is computationally
much more efficient. The approach proposed by Schut et al. (2021) and
used to generate the examples in Figure~\ref{fig-dynamics} falls into
the latter category. In future work, I want to explore how recent
advances in post-hoc uncertainty quantification, most notably Laplace
Redux (Daxberger et al. 2021), can be leveraged to generate realistic
and unambiguous counterfactual explanations for any model.\footnote{For
  some initial work on this see my Julia implementation of Laplace
  Redux:
  \href{https://www.paltmeyer.com/BayesLaplace.jl/dev/}{BayesLaplace.jl}.}

>>>>

\title{Dynamics in Algorithmic Recourse}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Trustworthy Artificial Intelligence for Finance and Economics}
\author{}
\date{April 29, 2022}

@@

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, sharp corners, frame hidden, borderline west={3pt}{0pt}{shadecolor}, interior hidden, enhanced, breakable]}{\end{tcolorbox}}\fi

@@

Recent advances in artificial intelligence (AI) have propelled its
adoption in domains outside of computer science including health care,
bioinformatics and genetics. In finance, economics and other social
sciences, applications of AI are still relatively limited.
Decision-making in these fields has traditionally been guided by
interpretable models that facilitate explanations. Explainability is
crucial in this context, since decision-makers are typically held
accountable by the public: central banks, for example, are heavily
scrutinized for the policies they impose. It is therefore not surprising
that practitioners and academics in these fields are reluctant to adopt
AI technologies they cannot trust. Deep neural networks, for example,
are generally considered as black boxes and therefore not trustworthy in
a context that demands explanations. This PhD project is focused on
exploring and developing methodologies that improve the trustworthiness
of AI and thereby enable its application in Finance and Economics.

The remainder of this extended abstract is structured as follows:
Section~\ref{sec-main} presents one of the research questions I have
investigated during the first months of my PhD: how do counterfactual
explanations handle dynamics? Section~\ref{sec-related} places this work
in the broader context of my research.

@@

\textbf{Counterfactual explanations} (CE) explain how inputs into a
model need to change for it to produce different outputs. They are
intuitive, simple and intrinsically linked to the potential outcome
framework for causal inference, which social scientists are familiar
with. Counterfactual explanations that involve realistic and actionable
changes can be used for the purpose of \textbf{Algorithmic Recourse}
(AR) to help individuals who face adverse outcomes. An example relevant
to the Finance and Economics domain is consumer credit: in this context
AR can be used to guide individuals in improving their creditworthiness,
should they have previously been denied access to credit based on an
automated decision-making system.

Existing work on CE and AR has largely been limited to the static
setting: given some classifier \(M: \mathcal{X} \mapsto \mathcal{Y}\) we
are interested in finding close (Wachter, Mittelstadt, and Russell
2017), actionable (Ustun, Spangher, and Liu 2019), realistic Schut et
al. (2021), sparse, diverse (Mothilal, Sharma, and Tan 2020) and ideally
causally founded counterfactual explanations (Karimi, Schölkopf, and
Valera 2021) for some individual \(x\). The ability of counterfactual
explanations to handle dynamics like data and model shifts remains a
largely unexplored research challenge at this point (Verma, Dickerson,
and Hines 2020). Only one recent work considers the implications of
\textbf{exogenous} domain and model shifts (Upadhyay, Joshi, and
Lakkaraju 2021). The authors propose a simple minimax objective, that
minimizes the counterfactual loss function for a maximal model shift.
They show that their approach yields more robust counterfactuals in this
context than existing approaches.

This project investigates \textbf{endogenous} domain and model shifts,
that is shifts that occur when AR is actually implemented by a
proportion of individuals and the classifier is updated in response.
Figure~\ref{fig-dynamics} illustrates this idea for a binary problem
involving a probabilistic classifier and the counterfactual generator
proposed by Wachter, Mittelstadt, and Russell (2017): the implementation
of AR for a subset of individuals leads to a domain shift (b), which in
turn triggers a model shift (c). As this game of implementing AR and
updating the classifier is repeated, the decision boundary moves away
from training samples that were originally in the target class (d).

These dynamics may be problematic. As the decision boundary moves in the
direction of the non-target class, counterfactual paths become shorter:
in the loan example, individuals that previously would have been denied
credit based on their input features are suddenly considered as
creditworthy. Average default risk across all borrowers can therefore be
expected to increase. Conversely, lenders that anticipate such dynamics
may choose to deny credit to individuals that have implemented AR,
thereby compromising the validity of AR.

To the best of my knowledge this is the first work investigating
endogenous dynamic in AR. Through future experiments I want to
investigate how this phenomenon plays out across different benchmark
datasets including German credit, Boston Housing and COMPAS.\footnote{These
  benchmark datasets have their issues and controversies, which is one
  of the challenges I would like to discuss at AIES.} Furthermore, I
want to assess to what extent the magnitude and direction of domain and
model shifts depends on the choice of the counterfactual generator. To
this end, I am currently supervising a group of undergraduate students,
who are tackling some of these tasks in their final-year research
project.

@@

\caption{\label{fig-dynamics}Dynamics in Algorithmic Recourse: we have a
simple Bayesian model trained for binary classification (a); the
implementation of AR for a random subset of individuals leads to a
domain shift (b); as the classifier is retrained we observe a model
shift (c); as this process is repeated, the decision boundary moves away
from the target class (d).}

@@

Until recently there existed only one open-source library that provides
a unifying approach to generate and benchmark counterfactual
explanations for Python models (Pawelczyk et al. 2021). To address this
limitation I have developed
\href{https://www.paltmeyer.com/CounterfactualExplanations.jl/stable/}{CounterfactualExplanations.jl}:
a Julia package that can be used to generate counterfactual explanations
for models developed and trained not only in Julia, but also in other
popular programming languages. The package and companion paper are
pending acceptance for a main talk at
\href{https://juliacon.org/2022/}{JuliaCon '22}.

@@

To ensure that the generated counterfactuals are realistic it helps to
understand which input-output pairs are likely to occur under the data
generating process. To this end, previous work has either relied on
generative models or restricted the analysis to probabilistic
classifiers that incorporate uncertainty in their predictions. While the
former approach is more generally applicable, the latter is
computationally more efficient. In future work, I want to explore how
recent advances in post-hoc uncertainty quantification, most notably
Laplace Redux (Daxberger et al. 2021), can be leveraged to generate
realistic and unambiguous counterfactual explanations for any
model.\footnote{For some initial work on this see my Julia
  implementation of Laplace Redux:
  \href{https://www.paltmeyer.com/BayesLaplace.jl/dev/}{BayesLaplace.jl}.}
With respect to the work-in-progress presented here, I expect that these
efforts may help in mitigating endogenous domain and model shifts.

### 2022-04-29 dev/submissions/aies/sections/cover.qmd
aies
Fun with latex templates
first complete draft

---
title: "Overlooked Dynamics in Algorithmic Recourse"
format: 
    html: default
    pdf:
        documentclass: acmart
        # include-in-header: preamble.tex
author:
  - name: Patrick Altmeyer
---

### Author

- Full Name: {{< var author.name >}}
- Home page: {{< var author.url >}}
- Affiliation: {{< var author.affiliation >}}
- Affiliation URL: {{< var author.affiliation-url >}}
- Country of citizenship: {{< var author.nationality >}}

### PhD Progress

- Expected graduation date: {{< var phd.graduation-date >}}
- Years in program: {{< var phd.years >}}

### Research Advisor

- Full Name: {{< var advisor.name >}}

@@

### Letter of Support

\pagebreak

## Extended abstract

>>>>

---
format:
    pdf: default
---

## Personal details

|  | Author | Research Advisor |
|---------|:--------|:--------|
| Name     | {{< var author.name >}} | {{< var advisor.name >}} |  
| Email | {{< var author.email >}} | {{< var advisor.email >}} | 
| Phone | {{< var author.tel >}} | {{< var advisor.tel >}} | 
| Home page | {{< var author.url >}} | {{< var advisor.url >}} | 
| Affiliation | {{< var author.affiliation >}} | {{< var advisor.affiliation >}} | 
| Affiliation URL | {{< var author.affiliation-url >}} | {{< var advisor.affiliation-url >}} | 
| Citizenship | {{< var author.nationality >}} | {{< var advisor.nationality >}} | 
| Expected graduation date | {{< var phd.graduation-date >}} | --- | 
| Years in program | {{< var phd.years >}} | --- |

@@

### 2022-04-29 dev/submissions/aies/sections/support.qmd
first complete draft

### Letter of Support

As substantiated in more detail in my full letter of support (attached with this application), as Patrick's advisor, I am very happy to support his application to the AIES student track. Patrick is an eager student, with a pro-active enthusiasm to learn from others, and contribute himself to the learning progress of relevant communities. Getting him to present at AIES will be very valuable for receiving feedback on his current work, while I would more broadly be happy to receive the community's guidance on how to best navigate questions of ethics and responsibility in a project like his. In case of a positive outcome of this application, we are in a position to fully cover Patrick's costs, and thus will not lay any claim on the financial waivers.

>>>>

## Letter of Support

> As substantiated in more detail in my full letter of support (attached with this application), as Patrick's advisor, I am very happy to support his application to the AIES student track. Patrick is an eager student, with a pro-active enthusiasm to learn from others, and contribute himself to the learning progress of relevant communities. Getting him to present at AIES will be very valuable for receiving feedback on his current work, while I would more broadly be happy to receive the community's guidance on how to best navigate questions of ethics and responsibility in a project like his. In case of a positive outcome of this application, we are in a position to fully cover Patrick's costs, and thus will not lay any claim on the financial waivers. 
> --- {{< var advisor.name >}}
